Functory: A Distributed Computing Library \\ for Objective Caml \thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2011-03-10 doc/tfp-2011/main.tex
updated introduction section [KK]

This paper introduces a generic library for distributed computing,
targeted at functional programming. 
This work was initially motivated by the computing needs that exist in our
own research team. Our applications include large-scale deductive
program verification, which amounts to checking the validity of a
large number of logical formulas using a variety of automated theorem
provers~\cite{filliatre07cav}. Our computing infrastructure consists
of a few powerful multi-core machines (typically 8 to 16 cores) and
several desktop PCs (typically dual-core). However, for our
application needs, there is no library that helps in exploiting such a
computing infrastructure in our favorite functional programming
language.  Hence we designed and implemented such a
library, which is the subject of this paper.
This library is implemented in \Ocaml, but the
implementation should be straightforward in any functional
programming language.

The distributed computing library presented in this paper is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of APIs that allows the
execution of large-scale parallelizable computations, very relevant to
our application needs (and actually relevant to a variety of
real-world applications), over multiple cores in the same machine or
over a network of machines. 
The most important features of our library are the following:
\begin{itemize}
\item \emph{Genericity}: 
  it allows various patterns of polymorphic computations;
\item \emph{Simplicity}: switching between multiple cores on the same
  machine and a network of machines is as simple as changing a couple
  of lines of code;
\item \emph{Task distribution and fault-tolerance}: 
  it provides automatic task distribution and
  a robust fault-tolerance mechanism, thereby
  relieving the user from implementing such routines.
\end{itemize}
The application domain of such a distributed computing library is manyfold. 
It is worth noting that the library is not targeted at applications
running on server farms, crunching enormous amounts of data, mostly
due to the lack of data locality. 
It is rather more relevant for research labs, to make use of available
computing resources efficiently. It serves a variety of users and a wide
spectrum of needs, from desktop PCs to networks of machines, and hence
the title.
Typical applications would involve a large number
of computation expensive tasks.
This is the case in our
research endeavours, that is validating thousands of verification
conditions using automated theorem provers.

>>>>

This paper introduces a generic library for distributed computing, for
a widely used functional programming language - OCaml. This work was
initially motivated by the computing needs that exist in our own
research team. Our applications include large-scale deductive program
verification, which amounts to checking the validity of a large number
of logical formulas using a variety of automated theorem
provers~\cite{filliatre07cav}. Our computing infrastructure consists
of a few powerful multi-core machines (typically 8 to 16 cores) and
several desktop PCs (typically dual-core). However, for our
application needs, there is no library that helps in exploiting such a
computing infrastructure in our favorite functional programming
language.  Hence we designed and implemented such a library, which is
the subject of this paper.  This library is implemented in \Ocaml, but
the implementation should be straightforward in any functional
programming language.

The distributed computing library presented in this paper is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of user-friendly APIs
that allows distributed execution of large-scale parallelizable
computations, very relevant to our application needs (and also
relevant to a variety of real-world applications). Further, the
distributed execution could be over multiple cores in the same machine
or over a network of machines. The most important features of our
library are the following:
\begin{itemize}
\item \emph{Genericity}: 
  it allows various patterns of polymorphic computations;
\item \emph{Simplicity}: switching between multiple cores on the same
  machine and a network of machines is as simple as changing a couple
  of lines of code;
\item \emph{Task distribution and fault-tolerance}: 
  it provides automatic task distribution and
  a robust fault-tolerance mechanism, thereby
  relieving the user from implementing such routines.
\end{itemize}
The application domain of such a distributed computing library is manyfold. 
% It is worth noting that the library is not targeted at applications
% running on server farms, crunching enormous amounts of data, mostly
% due to the lack of data locality. 
The library is very relevant for research labs, to make use of
available computing resources efficiently. It serves a variety of
users and a wide spectrum of needs, from desktop PCs to networks of
machines.  Typical applications would involve executing a large number
of computation expensive tasks in a resource-optimal and
time-efficient manner. This is also the case in our research
endeavours, that is validating thousands of verification conditions
using automated theorem provers, utilizing the computing
infrastructure to the maximum.
